---
title: "Disease columns"
author: "Sabrina Heitmann"
date: "5/12/2021"
output: html_document
---

Load packages
```{r}
library(tidyverse)
library(magrittr)
```

Tidy dataset, populates sex column, and populates a column with disease data to show that it can be done
```{r}
# upload fish data
zdata_formatted <- read.csv("final_formatted_data_v2.csv", na.strings=c("","NA"))

#preview dataset
glimpse(zdata_formatted)

# Pivots text longer (each fish has own row  )
zdata_tidy <- zdata_formatted %>%
  pivot_longer(
   cols = starts_with("Fish_Entry"),
   names_to = "fish",
   names_prefix = "Fish_Entry",
   values_to = "histo_inv_fish",
   values_drop_na = TRUE
 )

# keep only important columns (can adjust if needed)
zdata_reduced <- zdata_tidy %>%
  select(histo_inv_fish, CaseId, Histopathology, fish, CaseDescr_right)

# populates new sex column
filtered_sex <- zdata_reduced %>%
  filter(str_detect(histo_inv_fish, "[Mm]ale")) %>%
  group_by(CaseId) %>% mutate(fish = row_number()) %>%
  mutate(sex = str_extract(histo_inv_fish, "Female|Male")) %>%
  mutate(fish_id = paste(CaseId, "_", fish, sep = ""))

# Extracts disease from histopathology column and populated binary disease column (1 = presence, 0 = absence)
# This shows us that this can be done, however, we need to make a function so it can be automated for our list of diseases
disease_columns <- filtered_sex %>%
  mutate(Hepatic_meglocytosis = str_extract(histo_inv_fish, "[Hh]epatic megalocytosis")) %>%
  mutate(Hepatic_meglocytosis = if_else(Hepatic_meglocytosis == "[a-z]", 1, 1, 0))

```

1. Make a function that extracts a disease from each individual fish's histopathology and populates presence or absence in new column (binary)
```{r}
# function works except it doesn't name the column with the disease
disease_to_columns <- function(dataset, disease, column_name){
  dataset %>%
    mutate(column_name = str_extract(histo_inv_fish, disease)) %>%
    mutate(column_name = if_else(column_name == "[a-z]", 1, 1, 0))
}

filtered_wdisease <- disease_to_columns(filtered_sex, "[Hh]epatic megalocytosis", Hepatic_meglocytosis)
filtered_wdisease <- disease_to_columns(filtered_sex, "NSL", NSL)
```

2. Create a list of all diseases (and synonyms) we want to extract
```{r}


```

3. We're going to use nested for loops instead of lapply()
BR: this code block has not been tested for syntax or function; it's just what we ended with after meeting on 05.14.2021
next steps: find out what exactly mutate does, to know if we can just add it to this block below without redundancies.
```{r}

#filtered_wdisease
for (fish in filtered_wdisease$histo_inv_fish) {
  for(row_num in 1:nrow(synonyms)) {
    eachDisease_row = synonyms[row_num, ]
    #Create output column based on first column
    for (col_num in 1:nrow(synonyms))
      eachSynonym_column = synonyms[ , col_num]
      # if fish cell total string has smaller synonym string in it...
      # use mutate below?
      str_extract(fish, eachSynonym_column)
      # put the integer 1 in the output csv, in the column whose header name is the parent disease
      # the parent disease string is the object eachDisease_row[1] (i.e. the first element in the current row of synonyms)
      # use mutate below?
      disease_columns(fish, eachSynonym_column, eachDisease_row)
      # else, put the integer 0 in the column




```

4. Run pipeline on rows on already annotated disease columns and compare numbers
```{r}

```

5. Run pipeline on full dataset
```{r}

```

Outcome: Disease columns are now populated on a per fish basis
